Solar Flares Across the Spectrum: Observations and Instrumentation

### 2025-04-11 chapter3/chap3.tex
Local changes

There are several studies \citep{hilarie05,caspi10} that estimate the thermal energy as a function of time. The thermal energy of the flares is defined as 
\begin{equation}
\mathrm{U_{Th}~\simeq~3n_{e}k_{B}TV}f,
\end{equation} \label{eq:t_eneg_1}
 where `$\mathrm{n_{e}}$' is the electron number density of the flaring plasma, `V' is the volume of the flaring plasma, `\textit{f}' is the volume filling factor, and `T' is the instantaneous temperature. One of the key challenges in characterizing the flares' thermal energy is determining the volume.

@@

%%%%######%%%%%%
\begin{figure}[ht!]
    \centering
    \includegraphics[trim={0cm 0.5cm 1.2cm 0.5cm},clip,width=\textwidth]{flare_pos.pdf}
    \caption{Position of various instruments during the flares. AIA 193~{\AA} observations during 2021 Oct 28 (panel a) and EUVI 171~{\AA} observation during 2020 Nov 29 (panel c) in Heliographic Stonyhurst (HGS) coordinates. The two flares studied here are enclosed by white dotted circle. The projected positions of {\it Solar Orbiter}, {\it SDO} and {\it STEREO-A} are marked as labelled. Positions of various spacecrafts projected onto the Heliocentric Inertial (HCI) frame for the for 2021 Oct 28 (panel c) and 2020 Nov 29 (panel d), as labelled.}
    \label{fig:sc_pos}
\end{figure}
%%%%######%%%%%%

Here we study the evolution of the thermal energy of two flares, \textit{viz.} M-class and X-class. We use existing tools and observations of the same region on the Sun from different vantages to estimate the volume of the flaring plasma more accurately. We show that a better estimate of the plasma volume affects the estimate of the thermal energy and has implications towards the energy budget of the flare. We also estimate the cumulative non-thermal energy deposited by the non-thermal electrons and compare it with our thermal energy estimates for the flare observed on 28 October, 2021. In \S\ref{sec:obs} we discuss the observations used in this paper and details of our analysis. We discuss our method for estimating the thermal energy in \S\ref{sec:therm}. In \S\ref{sec:los} and \S\ref{sec:non-therm}, we describe our method of volume estimation and cumulative non-thermal energy estimation, respectively. We discuss the results in \S\ref{res}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{X flare on 28th October, 2021}\label{sec:x1-obs}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%--------------------------
\begin{figure}[ht!]
    %\hspace{-1cm}
    \centering
    \includegraphics[width=0.95\textwidth,trim={0.3cm 4cm 0cm 2cm},clip]{oct28_align_new.pdf}
    \caption{X-class flare observed on October 28, 2021. Panel (a): the post eruption arcade as observed in AIA 193~{\AA}. Panel (b): Two ribbon of the flare observed in AIA 1600 {\AA}, overplotted with STIX soft X-ray contours of 4{--}15 keV (solid blue lines) and hard X-ray contours of 15{--}25 keV (dashed red lines). Panel (c): the emission measure map of the region integrated over the temperature range of 5$<\log\,T<$7.4. Panel (d): the obtained DEM weighted temperature.}
    \label{fig:flare}
\end{figure}
%%--------------------------
The X1 flare occurred on Oct 28, 2021 (SOL2021-10-28T15:35) near $\sim~[0\arcsec,-500\arcsec]$ solar coordinates as seen from AIA. The flare was observed by {\it SDO}/AIA, {\it STEREO-A}/EUVI, {\it GOES}/SUVI, and XRT, along with {\it GOES}/XRS and STIX observations. AIA observations provide imaging with 94, 131, 193, 171, 211 and 335 {\AA} passbands. SUVI provides imaging with similar wavelengths as AIA: 94, 131, 195, 171, 195 and 284 {\AA}, while {\it STEREO-A}/EUVI provides imaging in 171, 195, 284 {\AA} coronal passbands. The flare exhibits a standard two-ribbon structure, along with a ridge at the top of the rising flare arcade \citep{longcope22}. The ridge most prominently shows up in the AIA 193 {\AA} channel with contributions from \ion{Fe}{12} and \ion{Fe}{24}. %\cite{longcope22} showed that the ridge was mainly created by the hotter \ion{Fe}{24} line (T~$\simeq$~17~MK) and a height of the flare arcade, $h~\simeq~20~Mm$. We use this ridge to identify the loop-top in the AIA observation and calculate the height of the loop (see \S\ref{sec:los} for further details).

@@

%%--------------------------
\begin{figure}[ht!]
\centering
    \includegraphics[trim={4.8cm 2.5cm 6cm 2.5cm},clip,width=0.8\textwidth]{nov_29_align.pdf}
    \caption{M-class flare on Nov 29, 2020 as observed nearly simultaneously by AIA~{131 \AA} (panel a) and XRT Be-med filer (panel b) and the derived EM map (panel c) and DEM weighted temperature (panel d). The EM map is obtained by integrating over the temperature range of 5$<\log\,T<$7.4.}
    \label{fig:flare2}
\end{figure}
%%--------------------------

@@

%%--------------
\begin{figure}[ht!]
    \centering
    \includegraphics[width=0.8\textwidth,trim={2cm 6cm 2cm 6cm},clip]{flare_orient.pdf}
    \caption{Panel (a): {\it STEREO-A}/EUVI 171 {\AA} observation of the flare arcade in the decay phase. The magenta cross marks a point on the top of the arcade. The line of sight (LOS) goes into the page through that point. Panel (b): SDO/AIA 171 {\AA} observation of the flare arcade. The red line marks the LOS through the arcade from the STEREO-A perspective projected to the AIA perspective.} %Panel (c): Cartoon demonstrating the projection effects of a semicircular loop between STEREO-A and AIA vantages due to the difference between polar ($\theta$) and azimuthal ($\phi$) angles respectively. Panel (d) : The calculated LOS map from the AIA perspective, with the calculated loop top height under the assumption of a semi-circular loop geometry.}
    \label{fig:flare_orient}
\end{figure}
%%--------------
%%--------------
\begin{figure}[ht!]
    \centering
    \includegraphics[width=0.9\textwidth,trim={4cm 5cm 3cm 5cm},clip]{nov29_triangulate.pdf}
    \caption{Panel (a): SDO/AIA 193 {\AA} observation of the Nov 29 flare in the decay phase. The blue cross marks a point on the top of the arcade. The LOS goes into the page from AIA perspective. Panel (b): {\it STEREO-A}/EUVI 195 {\AA} observation around same time. The blue solid line is the LOS from AIA perspective in panel (a) projected onto {\it STEREO-A} perspective. We can trace out the height of the flare arcade at various locations using \textit{scc\_measure.pro}. The height of the arcade at various points from the Sun's surface is marked in panel (a).}
    \label{fig:flare_orient2}
    \end{figure}
%%--------------

@@

%%--------------
\begin{figure}[ht!][!h]
    \centering
    \includegraphics[width=0.8\textwidth,trim={3cm 2cm 3cm 0cm},clip]{flare_orient_pt2.pdf}
    \caption{Panel (a): Cartoon demonstrating the projection effects of a semicircular loop between STEREO-A and AIA vantages due to the difference between polar ($\theta$) and azimuthal ($\phi$) angles, respectively. Panel (b) : The calculated LOS map from the AIA perspective for the 28th October, 2021 event, with the estimated loop top height under the assumption of a semi-circular loop geometry.}
    \label{fig:flare_orient_2}
\end{figure}
%%--------------

@@

To infer the LOS along individual pixels in the FOV for both the flares, we use AIA and STEREO-A/EUI observations, recorded from two different vantage points, to calculate the height of the flare loop. We display STEREO-A/EUI and AIA-171 image taken neeary simultaneously in Fig.~\ref{fig:flare_orient} a~\&~b, respectively. The magenta cross on Fig.~\ref{fig:flare_orient}.a shows a point at the top of the flare arcade. The LOS goes into the page through that point from the {\it STEREO-A} vantage. The red line in Fig.~\ref{fig:flare_orient}.b shows the LOS through the point in Fig.~\ref{fig:flare_orient} a projected to {\it SDO}/AIA point of view. This demonstrate the geometric effect of observing from different vantage points.

We show the same in Fig.~\ref{fig:flare_orient2} for the Nov 29, 2020 flare. The blue cross marked in Fig.~\ref{fig:flare_orient2}.a is the LOS going into the page from AIA perspective. The blue line marked in Fig.~\ref{fig:flare_orient2}.b is the same LOS projected onto the STEREO-A perspective. Distances on the observation from two different vantages would be projected by the differences in the polar and the azimuthal angle between the two vantages. Fig.~\ref{fig:flare_orient_2}.a shows the projection effect of a semicircular loop between AIA and STEREO-A LOS. Depending on the orientation of the loops, the projection along the polar and azimuthal angle projects the length of the loop and the height of the loop. The height at various position of the flare arcade is marked in Fig.~\ref{fig:flare_orient2}.a. We assume a semi-circular loop geometry with the height of the loop top to calculate the LOS along every pixel within the flare arcade.

We use co-temporal {\it STEREO-A} 171 {\AA}, 195 {\AA}, and AIA~171 {\AA}, 193 {\AA} images with \textit{scc\_measure.pro}, available in the \textit{sswidl}, to calculate the height of the loop. This routine allows us to select a point on the observation from one of the two vantages. The line of sight (LOS) through the same point is shown on the observation from the different vantage. The same point can be identified on the projected LOS by the user, by identifying similar emission characteristics. The program determines the 3D coordinates of the point (heliographic latitude, longitude and radial distance). We carry out the same measurement at various positions along the loop top, to calculate the change in loop height across the arcade. With the different height measurements at various locations, we can calculate the volume of the flaring plasma under a semicircular assumption. We note that the height estimation using \textit{scc\_measure.pro} is limited by our identification of ``similar emission characteristics" between the two observations from AIA and {\it STEREO-A}.

@@

Figure \ref{fig:flare_orient_2}.b shows an example of the calculated LOS in the FOV for the 2021 October 28 flare. The highest point of the flare arcade is oriented from south east towards north west, following the ridge. We get an estimate of the highest point of the loop $\simeq$ 22~Mm, which is consistent with the estimates of the height, $H~\simeq~20~Mm$ of the ridge from magnetic loop modelling by \cite{longcope22} for the same event. Using the map of the LOS shown in the bottom panel of Fig. \ref{fig:flare_orient_2}, we determine the thermal energy in every pixel of the flare using Eqn.~\ref{eq:t_eneg}.  A similar process is used for the flare loops in the 2020 November 29 flare.

@@

We estimate the energy in the non-thermal electrons for the 2021 Oct 28 X-class event by fitting the STIX spectra. In Fig.~\ref{fig:stix_an}.a, we plot the STIX light curve of the event in 4 to 6 keV (solid black line), 6 to 12 keV (solid magenta line), 12 to 25 keV (solid yellow line) and 25 to 50 keV (solid blue line). The attenuator kicked in around $\sim$ 15:28 UT, which drastically cuts down the intensity to avoid saturation. The hard X-ray peak is visible in the 25 to 50 keV (solid blue line) at $\sim$ 15:28 UT. The soft X-ray peak is visible within the attenuated flux in 6 to 12 keV (solid magenta line) at $\sim$ 15:30 UT.

%%--------------
\begin{figure}[ht!]
    \centering
    \includegraphics[trim={3cm 2cm 6cm 3.5cm},clip,width=0.54\textwidth]{oct28_lc_2.pdf} 
    \includegraphics[trim={3.5cm 8cm 3cm 2cm},clip,width=0.45\textwidth]{oct28_fit_2.pdf}
    \caption{Panel a: STIX light curve of the 2021 Oct 28 X-class event in different energy bands as labelled. Panel b: STIX spectra integrated between 15:27{--}15:28~UT during impulsive phase, fitted with `\textit{thcik2}' (solid green), `\textit{2vth}' (solid yellow) and the complete fit function `\textit{2vth+thick2}' (solid red line). The lower panel shows the normalized residuals of the fit.}
    \label{fig:stix_an}
    \end{figure}
%%--------------

We closely follow the method described in \cite{emslie12} to estimate the energy deposited by the non-thermal electrons. We fit the STIX spectra of the event averaged over several time bins during the evolution of the flare with `\textit{2vth}' and `\textit{thick2}' functions available in the `\textit{OSPEX}' X-ray spectra fitting package in {\it sswidl}. We show a reference fit to the spectra obtained over a time bin $\sim$ 15:26-15:27 UT in Fig.~\ref{fig:stix_an} panel (b) with `\textit{2vth}' (solid yellow) and `\textit{thick2}' (solid green). The function `\textit{2vth}' is a two-component thermal function. The parameters are the emission measure and temperature of the two thermal components and the relative abundances of Iron/Nickel, Calcium, Sulfur and Silicon with respect to the coronal abundances of CHIANTI \citep{chianti1,chianti}. The `\textit{thick2}' assumes the non-thermal component to be bremsstrahlung from energetic electrons with an injected spectrum $F_{0}(E_{0})~(e^{-}s^{-1}cm^{-2}keV^{-1})$ in the form of a broken power law:

@@

The parameters of this model spectrum, e.g., the normalization parameter A, the low- and high-energy cutoffs $E_{min}$ and $E_{max}$, break energy $E_{b}$, the power law indices $\delta_{1}$ and $\delta_{2}$ below and above the break are constrained by the fitting. We only fix the high energy cutoff at $3.2\times 10^{4}$ keV for every fit. The high energy cutoff is fixed much higher than our concerned energy range ($\sim$ O($10^{1}$ keV)), and we can safely assume that it has a negligible effect on fitting the X-ray spectra \citep{emslie12}.

The `\textit{OSPEX}' fitting is done by a forward fitting procedure. The parameters of the functions `\textit{2vth}' and `\textit{thick2}' are varied to generate a photon spectrum, which is folded through the detector response matrix to produce a count rate spectrum. This count rate spectrum is used to constrain the parameters of the functions through an iterative procedure by minimizing the $\chi^{2}$ between the calculated and measured count rate spectrum. We provide the fitted temperature and the EM of the hotter thermal component in Tab.~\ref{tab:tab1} during various phases of the flare. The temperatures listed in Tab.~\ref{tab:tab1} is lower than the upper limit of the temperature range used ($log(T_{max})~=~7.4$ and $log(T_{min})~=~5$) to calculate the EM and DEM weighted temperature in \S\ref{sec:therm}.

%%%%%%%%%%%
\begin{table}
    \centering
    \resizebox{0.9\textwidth}{!}{%
    \begin{tabular}{|cl||c|c|}
    \hline
         &  & log(T) & EM \\ 
         & \hspace{-1cm}Time (UT) & of hotter component & ($cm^{-3}$) \\
    \hline
        Impulsive phase & \hspace{1.5cm}15:26 & 7.098 & $10^{48}$\\
        HXR peak & $\sim$ 15:27:20 - 15:28:20 & 7.19 & $3\times 10^{48}$\\
        Right after SXR peak & $\sim$ 15:35:39 - 15:36:34 & 6.88 & $4.1\times 10^{49}$\\
        Into the decay phase & \hspace{0.9cm}14:48:20 & 6.5 & $2\times 10^{49}$\\
        \hline
    \end{tabular}}
    \caption{Fitted temperature of the hotter component of the thermal plasma during various stages of the flare.}
    \label{tab:tab1}
\end{table}
%%%%%%%%%%%

The non-thermal energy in the electrons ($U_{e}$) at any instant $t=t'$ can be estimated by integrating the best-fit electron energy spectrum via: $$U_{e}(t=t')~=~\Delta t\int_{E_{min}}^{E_{max}}~F_{0}(E_{0})(t=t')dE_{0}$$ where the electron energy distribution in constrained by fitting the count spectra for the time interval $t'-\Delta t/2 \le t \le t'+\Delta t/2$. The cumulative energy deposited into the foot point over time by the non-thermal electrons is a good indicator for the source of the thermal energy of the plasma. We add the energy deposited at the foot points by the non-thermal electrons from every time bin, to estimate the cumulative energy deposited by the non-thermal electrons up until that instant: $$U^{cumulative}_{e}(t)~=~\sum_{t'=t_{0}}^{t}U_{e}(t')$$

@@

%%###########%%
\begin{figure}[ht!]
    \centering
    \includegraphics[trim={4cm 1cm 4cm 1cm},clip,width=0.8\textwidth]{flare_eneg.pdf}
    \caption{Panel (a): Temporal evolution of thermal energy for the 2021 Oct 28 flare as measured using GOES light-curve using the effective volume and a filling factor \textit{f}~=~1 (solid red), \textit{f}~=~0.4 (green triangles), \textit{f}~=~0.2 (blue squares), thermal energy calculated from GOES light curve, but the volume inferred from the area of the of the flaring arcade with STIX SXR images, under the assumption that $V\sim A^{\frac{3}{2}}$ (purple dotted), and DEM obtained from AIA observations (black crosses). We also plot the evolution of non-thermal energy as estimated from STIX (lime green circles). Panel (b): Temporal evolution of thermal for the 2020 Nov 29 flare as measured from GOES light curves with a constant effective volume (solid red) and that calculated from the DEMs inferred from the imaging and a varying volume from the imaging (solid blue).}
    \label{fig:eneg}
\end{figure}
%%###########%%

>>>>

There are several studies \citep{hilarie05,caspi10} that estimate thermal energy as a function of time. The thermal energy of the flares is defined as 
\begin{equation}
\mathrm{U_{Th}~\simeq~3n_{e}k_{B}TV}f,
\end{equation} \label{eq:t_eneg_1}
 where `$\mathrm{n_{e}}$' is the electron number density of the flaring plasma, `V' is the volume of the flaring plasma, `\textit{f}' is the volume filling factor, and `T' is the instantaneous temperature. One of the key challenges in characterizing the flares' thermal energy is determining the volume.

@@

%%%%######%%%%%%
\begin{figure}[ht!]
    \centering
    \includegraphics[trim={0cm 0.5cm 1.2cm 0.5cm},clip,width=\textwidth]{flare_pos.pdf}
    \caption[Position of various instruments during the flares.]{Position of various instruments during the flares. AIA 193~{\AA} observations during 2021 Oct 28 (panel a) and EUVI 171~{\AA} observation during 2020 Nov 29 (panel c) in Heliographic Stonyhurst (HGS) coordinates. The two flares studied here are enclosed by white dotted circle. The projected positions of {\it Solar Orbiter}, {\it SDO} and {\it STEREO-A} are marked as labelled. Positions of various spacecrafts projected onto the Heliocentric Inertial (HCI) frame for the for 2021 Oct 28 (panel c) and 2020 Nov 29 (panel d), as labelled.}
    \label{fig:sc_pos}
\end{figure}
%%%%######%%%%%%

Here we study the evolution of the thermal energy of two flares, \textit{viz.} M-class and X-class. We use existing tools and observations of the same region on the Sun from different vantages to estimate the volume of the flaring plasma more accurately. We show that a better estimate of the plasma volume affects the estimate of the thermal energy and has implications towards the energy budget of the flare. We also estimate the cumulative non-thermal energy deposited by the non-thermal electrons and compare it with our thermal energy estimates for the flare observed on 28 October, 2021. In \S\ref{sec:obs} we discuss the observations used in this {\bf chapter} and details of our analysis. We discuss our method for estimating the thermal energy in \S\ref{sec:therm}. In \S\ref{sec:los} and \S\ref{sec:non-therm}, we describe our method of volume estimation and cumulative non-thermal energy estimation, respectively. We discuss the results in \S\ref{res}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{X flare on 28th October, 2021}\label{sec:x1-obs}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%--------------------------
\begin{figure}[ht!]
    %\hspace{-1cm}
    \centering
    \includegraphics[width=0.95\textwidth,trim={0.3cm 4cm 0cm 2cm},clip]{oct28_align_new.pdf}
    \caption[Observation of the X-class flare on October 28, 2021.]{X-class flare observed on October 28, 2021. Panel (a): the post eruption arcade as observed in AIA 193~{\AA}. Panel (b): Two ribbon of the flare observed in AIA 1600 {\AA}, overplotted with STIX soft X-ray contours of 4{--}15 keV (solid blue lines) and hard X-ray contours of 15{--}25 keV (dashed red lines). Panel (c): the emission measure map of the region integrated over the temperature range of 5$<\log\,T<$7.4. Panel (d): the obtained DEM weighted temperature.}
    \label{fig:flare}
\end{figure}
%%--------------------------
The X1 flare occurred on Oct 28, 2021 (SOL2021-10-28T15:35) near $\sim~[0\arcsec,-500\arcsec]$ solar coordinates as seen from AIA. The flare was observed by {\it SDO}/AIA, {\it STEREO-A}/EUVI, {\it GOES}/SUVI, and XRT, along with {\it GOES}/XRS and STIX observations. AIA observations provide imaging with 94, 131, 193, 171, 211 and 335 {\AA} passbands. SUVI provides imaging with similar wavelengths as AIA: 94, 131, 195, 171, 195 and 284 {\AA}, while {\it STEREO-A}/EUVI provides imaging in 171, 195, 284 {\AA} coronal passbands. The flare exhibits a standard two-ribbon structure, along with a ridge at the top of the rising flare arcade \citep{longcope22}. The ridge most prominently shows up in the AIA 193 {\AA} channel with contributions from \ion{Fe}{12} and \ion{Fe}{24}. %\cite{longcope22} showed that the ridge was mainly created by the hotter \ion{Fe}{24} line (T~$\simeq$~17~MK) and a height of the flare arcade, $h~\simeq~20~Mm$. We use this ridge to identify the loop-top in the AIA observation and calculate the height of the loop (see \S\ref{sec:los} for further details).

@@

%%--------------------------
\begin{figure}[ht!]
\centering
    \includegraphics[trim={4.8cm 2.5cm 6cm 2.5cm},clip,width=0.8\textwidth]{nov_29_align.pdf}
    \caption[Observation of the M-class flare on November 29, 2020.]{M-class flare on Nov 29, 2020 as observed nearly simultaneously by AIA~{131 \AA} (panel a) and XRT Be-med filer (panel b) and the derived EM map (panel c) and DEM weighted temperature (panel d). The EM map is obtained by integrating over the temperature range of 5$<\log\,T<$7.4.}
    \label{fig:flare2}
\end{figure}
%%--------------------------

@@

@@

@@

We use co-temporal {\it STEREO-A} 171 {\AA}, 195 {\AA}, and AIA~171 {\AA}, 193 {\AA} images with \textit{scc\_measure.pro}, available in the \textit{sswidl}, to calculate the height of the loop. This routine allows us to select a point on the observation from one of the two vantages. The line of sight (LOS) through the same point is shown on the observation from the different vantage. The same point can be identified on the projected LOS by the user by identifying similar emission characteristics. The program determines the 3D coordinates of the point (heliographic latitude, longitude and radial distance). We carry out the same measurement at various positions along the loop top, to calculate the change in loop height across the arcade. With the different height measurements at various locations, we can calculate the volume of the flaring plasma under a semicircular assumption. We note that the height estimation using \textit{scc\_measure.pro} is limited by our identification of ``similar emission characteristics" between the two observations from AIA and {\it STEREO-A}. We choose multiple points on the top of the arcade, and calculate the height of the arcade at various locations tracing the looptop. The number of points we consider is not fixed and varies from frame to frame. After this, using the chosen points we interpolate to get the loop top from the AIA perspective, and the calculated height at these points gives us how the height of the loop top changes across the arcade. We assume a semi-circular loop geometry with the height of the loop top to calculate the LOS along every pixel within the flare arcade. It is worth reiterating that a majority of the uncertainty in determining the LoS would be arising from the identification of the loop top in two different vantages.

To infer the LoS along individual pixels in the FOV for both the flares, we use AIA and STEREO-A/EUI observations, recorded from two different vantage points, to calculate the height of {\bf both flare loops}. We show the STEREO-A/EUI 195~{\AA} and AIA 193~{\AA} observation for the Nov 29, 2020 flare in Fig.~\ref{fig:flare_orient2} panel b and a respectively. The blue cross marked in Fig.~\ref{fig:flare_orient2}.a is the LOS going into the page from AIA perspective. The blue line marked in Fig.~\ref{fig:flare_orient2}.b is the same LOS projected onto the STEREO-A perspective. The red crosses marked on the AIA 193~{\AA} loops mark the LoS going into the page, to infer the looptop from the AIA perspective. The red crosses marked on Fig.~\ref{fig:flare_orient2}.b EUVI 195~{\AA} observations marks the position of the looptop along the LoS from AIA perspective considered to find the looptop from EUVI perspective. In Fig.~\ref{fig:flare_orient2}.c we show the inferred LoS map for these two frames. The fan is assumed to have a constant LoS depth as mentioned earlier. For the loops we calculate the LoS assuming a semi-circular loop geometry with the height inferred from \textit{scc\_measure.pro}. The inferred LoS is lower near the loop top, and increases as we move closer the the foot points, exhibiting the variable LoS across the observation frame.

We display STEREO-A/EUVI and AIA-171 {\bf images} taken {\bf nearly} simultaneously in Fig.~\ref{fig:flare_orient} a~\&~b, respectively for the 28 Oct, 2021 event. The red and black cross on Fig.~\ref{fig:flare_orient}.a shows a point at the top of the flare arcade. The LOS goes into the page through that point from the {\it STEREO-A} vantage. The blue line in Fig.~\ref{fig:flare_orient}.b shows the LOS through the point in Fig.~\ref{fig:flare_orient} a projected to {\it SDO}/AIA point of view. This demonstrates the geometric effect of observing from different vantage points. In Fig.~\ref{fig:flare_orient}.c we show the inferred LoS map from the AIA perspective.

%We show the same in Fig.~\ref{fig:flare_orient2} for the Nov 29, 2020 flare. The blue cross marked in Fig.~\ref{fig:flare_orient2}.a is the LOS going into the page from AIA perspective. The blue line marked in Fig.~\ref{fig:flare_orient2}.b is the same LOS projected onto the STEREO-A perspective. The red crosses marked on the AIA 193~{\AA} loops mark the LoS going into the page, to infer the looptop from the AIA perspective. The red crosses marked on Fig.~\ref{fig:flare_orient2}.b EUVI 195~{\AA} observations marks the position of the looptop along the LoS from AIA perspective considered to find the looptop from EUVI perspective. Distances on the observation from two different vantages would be projected by the differences in the polar and the azimuthal angle between the two vantages.

%Fig.~\ref{fig:flare_orient_2}.a shows the projection effect of a semicircular loop between AIA and STEREO-A LOS. Depending on the orientation of the loops, the projection along the polar and azimuthal angle projects the length of the loop and the height of the loop. The height at various {\bf positions} of the flare arcade is marked in Fig.~\ref{fig:flare_orient2}.a. We assume a semi-circular loop geometry with the height of the loop top to calculate the LOS along every pixel within the flare arcade.

%%--------------
\begin{figure*}[!h]
    \centering
    \includegraphics[width=\textwidth, trim={1cm 0cm 1cm 3cm}, clip]{paper_plot_los.pdf}
    \caption[LoS triangulation for the Nov 29th, 2020 flare]{(a) SDO/AIA 193 {\AA} observation of the Nov 29 flare in the decay phase. The red and blue crosses marks various points on the top of the arcade used to trace the looptop. The LOS goes into the page from AIA perspective. (b) {\it STEREO-A}/EUVI 195 {\AA} observation around same time. The blue solid line is the LOS from AIA perspective in panel (a) projected onto {\it STEREO-A} perspective. The red crosses mark the other red crosses from AIA perspective marked on the looptop visible from Stereo-A perspective. The line of sight of the region is shown in panel (c). The LoS is higher near the base of the loops and lower near the looptop. The fan is assumed to be constant LoS.}
    \label{fig:flare_orient2}
\end{figure*}
%%--------------

%%--------------
\begin{figure*}
    \centering
    \includegraphics[width=\textwidth, trim={1cm 0cm 1cm 3cm}, clip]{paper_plot_los_2.pdf}
    \caption[LoS triangulation for the Oct 28th, 2021 flare]{(a) {\it STEREO-A}/EUVI 171 {\AA} observation of the {\bf Oct 28} flare arcade in the decay phase. The red and black cross marks a point on the top of the arcade. The line of sight (LOS) goes into the page through that point. (b) SDO/AIA 171 {\AA} observation of the flare arcade. The blue line marks the LOS through the arcade from the STEREO-A perspective projected to the AIA perspective. The red crosses mark the other red crosses from Stereo-A perspective marked on the loop top visible from AIA perspective. The line of sight of the region is shown in panel (c).}
    \label{fig:flare_orient}
\end{figure*}
%%--------------

@@

@@

We estimate the energy in the non-thermal electrons for the 2021 Oct 28 X-class event by fitting the STIX spectra. In Fig.~\ref{fig:stix_an}.a, we plot the STIX light curve of the event in 4 to 10 keV (solid blue line), 10 to 15 keV (solid orange line), 15 to 25 keV (solid green line) and 25 to 50 keV (solid red line). The attenuator kicked in around $\sim$ 15:28 UT, which drastically cuts down the intensity to avoid saturation. The hard X-ray peak is visible in the 25 to 50 keV (solid blue line) at $\sim$ 15:28 UT. The soft X-ray peak is visible within the attenuated flux in 6 to 12 keV (solid magenta line) at $\sim$ 15:30 UT.

%%--------------
\begin{figure*}
    \centering
    \includegraphics[trim={0cm 3cm 1.3cm 3cm}, clip, width=\textwidth]{Figures/oct_28_lc_3.pdf}
    \caption[STIX light curve and spectra fit for the 2021 Oct 28 flare]{Panel a: STIX light curve of the 2021 Oct 28 X-class event in different energy bands as labeled. Panel b: STIX spectra integrated between 15:27:00{--}15:27:10~UT during impulsive phase, fitted with `\textit{thcik2}' (solid yellow), `\textit{vth}' (solid green) and the complete fit function `\textit{vth+thick2}' (solid red line). The panel (c) shows the normalized residuals of the fit.}
    \label{fig:stix_an}
    \end{figure*}
%%--------------

We closely follow the method described in \cite{emslie12} to estimate the energy deposited by the non-thermal electrons. We fit the STIX spectra of the event averaged over several time bins during the evolution of the flare with `\textit{vth}' and `\textit{thick2}' functions available in the `\textit{OSPEX}' X-ray spectra fitting package in {\it sswidl}. We show a reference fit to the spectra obtained over a time bin $\sim$ 15:27:00-15:27:10 UT in Fig.~\ref{fig:stix_an} panel (b) with `\textit{vth}' (solid green) and `\textit{thick2}' (solid yellow). The function `\textit{vth}' is a two-component thermal function. The parameters are the emission measure and temperature of the two thermal components and the relative abundances of Iron/Nickel, Calcium, Sulfur and Silicon with respect to the coronal abundances of CHIANTI \citep{chianti1,chianti}. The `\textit{thick2}' assumes the non-thermal component to be bremsstrahlung from energetic electrons with an injected spectrum $F_{0}(E_{0})~(e^{-}s^{-1}cm^{-2}keV^{-1})$ in the form of a broken power law:

@@

The parameters of this model spectrum, e.g., the normalization parameter A, the low- and high-energy cutoffs $E_{min}$ and $E_{max}$, break energy $E_{b}$, the power law indices $\delta_{1}$ and $\delta_{2}$ below and above the break are constrained by the fitting. We only fix the high energy cutoff at $3.2\times 10^{4}$ keV for every fit. The high energy cutoff is fixed much higher than our concerned energy range ($\sim$ O($10^{1}$ keV)), and we can safely assume that it has a negligible effect on fitting the X-ray spectra \citep{emslie12}. {\bf While this serves our purpose of quantifying the bulk non-thermal energy as a function of time, there can be finer details of the physical parameters from the "cold thick target" model that can pose non-physical conditions during the early rise phase of the flares. This can also have noticeable effect on the total thermal energy budget of the flare. To mitigate this several modifications of the cold thick target model have been proposed, e.g., local re-acceleration of electrons and ions \citep{brown09}, warm thick target model \citep{kontar15, kontar19} and injection of kappa distributed electrons instead of power law \citep{kasparova09, bataglia15, effenberger17}. We describe the physical origin of some of these models and their effects on the inferred injected electron spectra in Appendix~\ref{c:a2}. We also fit two of the modified models to the same spectra and compare the differences that arise due to that.}

The `\textit{OSPEX}' fitting is done by a forward fitting procedure. The parameters of the functions `\textit{vth}' and `\textit{thick2}' are varied to generate a photon spectrum, which is folded through the detector response matrix to produce a count rate spectrum. This count rate spectrum is used to constrain the parameters of the functions through an iterative procedure by minimizing the $\chi^{2}$ between the calculated and measured count rate spectrum. We provide the fitted temperature and the EM of the hotter thermal component in Tab.~\ref{tab:tab1} during various phases of the flare. The temperatures listed in Tab.~\ref{tab:tab1} is lower than the upper limit of the temperature range used ($5.0 \geq \log\,T \leq 7.4)$ %($log(T_{max})~=~7.4$ and $log(T_{min})~=~5$) 
to calculate the EM and DEM weighted temperature in \S\ref{sec:therm}.

%%%%%%%%%%%
\begin{table}
    \centering
    \resizebox{0.9\textwidth}{!}{%
    \begin{tabular}{|cl||c|c|}
    \hline
         &  & log(T) & EM \\ 
         & \hspace{-1cm}Time (UT) & of thermal component & ($cm^{-3}$) \\
    \hline
        Impulsive phase & \hspace{1.5cm}15:26 & 7.098 & $10^{48}$\\
        HXR peak & $\sim$ 15:27:20 - 15:28:20 & 7.19 & $3\times 10^{48}$\\
        Right after SXR peak & $\sim$ 15:35:39 - 15:36:34 & 6.88 & $4.1\times 10^{49}$\\
        Into the decay phase & \hspace{0.9cm}14:48:20 & 6.5 & $2\times 10^{49}$\\
        \hline
    \end{tabular}}
    \caption{Fitted temperature of the hotter component of the thermal plasma during various stages of the flare.}
    \label{tab:tab1}
\end{table}
%%%%%%%%%%%

The non-thermal energy in the electrons ($U_{e}$) at any instant $t=t'$ can be estimated by integrating the best-fit electron energy spectrum via: $$U_{e}(t=t')~=~\Delta t\int_{E_{min}}^{E_{max}}~F_{0}(E_{0})(t=t')dE_{0}$$ where the electron energy distribution in constrained by fitting the count spectra for the time interval $t'-\Delta t/2 \le t \le t'+\Delta t/2$. The cumulative energy deposited into the foot point over time by the non-thermal electrons is a good indicator for the source of the thermal energy of the plasma. We add the energy deposited at the foot points by the non-thermal electrons from every time bin, to estimate the cumulative energy deposited by the non-thermal electrons up until that instant: $$U^{cumulative}_{e}(t)~=~\sum_{t'=t_{0}}^{t}U_{e}(t')$$.

@@

%%###########%%
\begin{figure}[ht!]
    \centering
    \includegraphics[trim={4cm 1cm 4cm 1cm},clip,width=0.8\textwidth]{flare_eneg.pdf}
    \caption[Temporal evolution of the thermal and non-thermal energy for both flares.]{Panel (a): Temporal evolution of thermal energy for the 2021 Oct 28 flare as measured using GOES light-curve using the effective volume and a filling factor \textit{f}~=~1 (solid red), \textit{f}~=~0.4 (green triangles), \textit{f}~=~0.2 (blue squares), thermal energy calculated from GOES light curve, but the volume inferred from the area of the of the flaring arcade with STIX SXR images, under the assumption that $V\sim A^{\frac{3}{2}}$ (purple dotted), and DEM obtained from AIA observations (black crosses). We also plot the evolution of non-thermal energy as estimated from STIX (lime green circles). Panel (b): Temporal evolution of thermal for the 2020 Nov 29 flare as measured from GOES light curves with a constant effective volume (solid red) and that calculated from the DEMs inferred from the imaging and a varying volume from the imaging (solid blue).}
    \label{fig:eneg}
\end{figure}
%%###########%%
